---
title: "Floor analysis"
author: "Andrea"
date: "2022-11-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo = FALSE}
library('dplyr')
library('car')
library('brms')
library('purrr')
library('tidyr')
library('ggplot2')
library('ggridges')
library('viridis')
```

## Floor analysis

# Load data
```{r}
d_sub <- read.csv('/Users/addena/Documents/R/LSTM/Sand_vs_leaflitter_August.csv', 
                  sep=",") # substrate experiment
d_nat <- read.csv('/Users/addena/Documents/R/LSTM/naturalistic_exp.csv', 
                  sep=",") # naturalistic experiment
d_pent <- read.csv('/Users/addena/Documents/R/LSTM/C15_vs_PO_data.csv', 
                  sep=",") # pentadecane experiment
```

# Re-organise data to standardise column names and variables

```{r}
d_sub <- rename(d_sub, stim.out=leaf.out)
d_sub <- rename(d_sub, con.out=sand.out)
d_sub <- rename(d_sub, stim.right=sand.right)
d_sub <- rename(d_sub, stim=leaf)
d_sub <- rename(d_sub, con=sand)
d_sub$dayday <- ifelse(d_sub$day==1, 'one', 'two')
d_sub$stim.right[d_sub$stim.right==1] <- 'con'
d_sub$stim.right[d_sub$stim.right==-1] <- 'stim'
d_sub$condition <- rep('substrate', length(d_sub$date))
d_pent <- rename(d_pent, stim.right=stim.blue)
d_pent$stim.right[d_pent$stim.right==1] <- 'stim'
d_pent$stim.right[d_pent$stim.right==-1] <- 'con'
d_pent$female.age.days.oldest <- rep(NA, length(d_pent$date))
d_pent$dayday <- ifelse(d_pent$day==1, 'one', 'two')
d_pent$condition <- rep('pentadecane', length(d_pent$date))
d_nat$dayday <- ifelse(d_nat$day==1, 'one', 'two')
d_nat$condition <- rep('naturalistic', length(d_nat$date))
```

We are *not* going to aggregate the data by day, as we would normally do,
because in this case we are also interested in how the location in the cage
(side) affects the outcome, and since that changes every day, we cannot sum
day 1 and day 2.

Number of larvae deposited on each day can be quite different - sometimes we get
the majority of larvae on day 1 and almost nothing on day 2, or the other way
around. This is why we won't do the analysis using larval numbers directly, but
we will calculate the preference index instead. A PI of 1 indicates here that 
all flies larviposited in the stimulus trays (any tray), and a PI of -1 
indicates that all flies larviposited on the floor (any side).

We also need to correct the preference index for total tray vs. floor area.
The total floor surface area is much bigger than total tray surface area. We
will therefore calculate the preference index not based on the number of larvae,
but on the density of larvae (larvae per cm^2): 

In the 'pentadecane' experiment, the total floor area was 56.5*87 cm = 4915.5 cm^2,
subdivided into tray area = 1960 cm^2 and floor area = 2955.5 cm^2. 

In the 'substrate' and 'naturalistic' experiments, the total floor area was
56.5*56.5 cm = 3192.25 cm^2, subdivided into tray area = 980 cm^2 and floor area =
2212.25 cm^2.

```{r}
# 'pentadecane' condition
d_pent$tray_percm2 <- (d_pent$stim + d_pent$con) / 1960 # tray pupae per cm2
d_pent$floor_percm2 <- (d_pent$stim.out + d_pent$con.out) / 2955.5 # floor pupae per cm2
d_pent$PIdens <- (d_pent$tray_percm2 - d_pent$floor_percm2) /
                  (d_pent$tray_percm2 + d_pent$floor_percm2)
d_pent

# 'substrate' condition
d_sub$tray_percm2 <- (d_sub$stim + d_sub$con) / 980
d_sub$floor_percm2 <- (d_sub$stim.out + d_sub$con.out) / 2212.25
d_sub$PIdens <- (d_sub$tray_percm2 - d_sub$floor_percm2) /
                  (d_sub$tray_percm2 + d_sub$floor_percm2)
d_sub

# 'naturalistic' condition
d_nat$tray_percm2 <- (d_nat$stim + d_nat$con) / 980
d_nat$floor_percm2 <- (d_nat$stim.out + d_nat$con.out) / 2212.25
d_nat$PIdens <- (d_nat$tray_percm2 - d_nat$floor_percm2) /
                (d_nat$tray_percm2 + d_nat$floor_percm2)
d_nat

d_all <- rbind(d_sub, d_pent, d_nat)
d_all
```
# Overall analysis

```{r}
Anova(aov(PIdens ~ condition + dayday + stim.right, data = d_all))
```

Anova suggests that the preference for tray/floor depends on significantly on
the day of the experiment, and very weakly on experimental condition.

Let's look at the 'day' analysis first.

```{r}
ggplot(d_all, aes(x = dayday, y = PIdens, fill = condition)) +
  geom_boxplot(notch = T, notchwidth = 0.9) +
  scale_fill_viridis(discrete = T, alpha = 0.6) +
  ylim(-1,1) +
  coord_flip() +
  theme_classic() + 
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey')
```
OK! So we get more floor depositions on day 2, a LOT more in the case of the
'substrate' and 'pentadecane' condition but there is even a trend in the
'naturalistic' condition. Let's test the distributions against each other.

```{r}
ks.test(subset(d_all, condition=='substrate' & dayday=='one')$PIdens, 
        subset(d_all, condition=='substrate' & dayday=='two')$PIdens)
ks.test(subset(d_all, condition=='naturalistic' & dayday=='one')$PIdens, 
        subset(d_all, condition=='naturalistic' & dayday=='two')$PIdens)
ks.test(subset(d_all, condition=='pentadecane' & dayday=='one')$PIdens, 
        subset(d_all, condition=='pentadecane' & dayday=='two')$PIdens)

wilcox.test(subset(d_all, condition=='substrate' & dayday=='one')$PIdens, 
        subset(d_all, condition=='substrate' & dayday=='two')$PIdens)
wilcox.test(subset(d_all, condition=='naturalistic' & dayday=='one')$PIdens, 
        subset(d_all, condition=='naturalistic' & dayday=='two')$PIdens)
wilcox.test(subset(d_all, condition=='pentadecane' & dayday=='one')$PIdens, 
        subset(d_all, condition=='pentadecane' & dayday=='two')$PIdens)

```
Interesting. Only the Wilcoxon rank sum test shows a significant difference in 
the median, and only for the 'substrate' condition. The Kolmogorov-Smirnov test
does not think the distributions are different (although some come close).

Let's see if we see something if we bunch all conditions together.

```{r}
ks.test(subset(d_all, dayday=='one')$PIdens, 
        subset(d_all, dayday=='two')$PIdens)
wilcox.test(subset(d_all, dayday=='one')$PIdens, 
        subset(d_all, dayday=='two')$PIdens)
```

Interesting. So we do see a significant trend towards larviposition on the floor
on day 2 when we consider all experimental conditions together. Let's plot that.

```{r}
ggplot(d_all, aes(x = dayday, y = PIdens, fill = dayday)) +
  geom_boxplot(notch = T, notchwidth = 0.9) +
  scale_fill_viridis(discrete = T, alpha = 0.6) +
  ylim(-1,1) +
  coord_flip() +
  theme_classic() + 
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey')
```
We can see from the notches that these distributions are significantly different.
Interesting. I think this is result number 1 for figure S2. In this vein, we
will definitely want a plot of overall density PIs for the three conditions. 
But we need overall values for that instead of by-day, so let's calculate a bit 
more.

```{r}
d_all_cage <- aggregate(cbind(con, con.out, stim, stim.out) ~ cage + condition,
                        data= d_all, FUN = sum, na.rm = T)
d_all_cage$tray_percm2 <- (d_all_cage$stim + d_all_cage$con) / 1960 # tray pupae per cm2
d_all_cage$floor_percm2 <- (d_all_cage$stim.out + d_all_cage$con.out) / 3440 # floor pupae per cm2
d_all_cage$PIdens <- (d_all_cage$tray_percm2 - d_all_cage$floor_percm2) /
                  (d_all_cage$tray_percm2 + d_all_cage$floor_percm2)
d_all_cage
```

Now plot:

```{r}
ggplot(d_all_cage, aes(x = condition, y = PIdens, fill = condition)) +
  geom_boxplot(notch = T, notchwidth = 0.9) +
  geom_jitter(color = 'black', width = 0.2, size = 0.4) +
  scale_fill_viridis(discrete = T, alpha = 0.6) +
  ylim(-1,1) +
  coord_flip() +
  theme_classic() + 
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey')
```
Ok looks good, the corrected PIs definitely cluster above 0, that's good news.

# Special case - 'pentadecane' experiment

We know from a previous analysis that stimulus side *does* have an effect on 
larvipositions on the floor but not on the trays. Let's quickly analyse this,
too.

First, let's calculate actual preference indices, for tray and floor. These do 
not need to be corrected.

```{r}
d_all$PItray <- (d_all$stim - d_all$con) / (d_all$stim + d_all$con)
d_all$PIfloor <- (d_all$stim.out - d_all$con.out) / (d_all$stim.out + d_all$con.out)
d_all
```
Another way of showing this, because the plot of PIfloor by stim.right is
confusing... Let's calculate a preference index for side

```{r}
d_all$PIside_tray <- ifelse(d_all$stim.right=='stim', (d_all$stim - d_all$con)/
                                                         (d_all$stim + d_all$con),
                    (d_all$con - d_all$stim)/(d_all$con + d_all$stim))
d_all$PIside_floor <- ifelse(d_all$stim.right=='stim', (d_all$stim.out - d_all$con.out)/
                                                         (d_all$stim.out + d_all$con.out),
                    (d_all$con.out - d_all$stim.out)/(d_all$con.out + d_all$stim.out))
```

Now stats and plots.

```{r}
Anova(aov(PIfloor ~ stim.right + dayday, data=subset(d_all, condition=='pentadecane'))) # very strong effect of side
ks.test(subset(d_all, condition=='pentadecane')$PIside_tray,
        subset(d_all, condition=='pentadecane')$PIside_floor) # p = 0.00339
wilcox.test(subset(d_all, condition=='pentadecane')$PIside_tray,
        subset(d_all, condition=='pentadecane')$PIside_floor) # p = 0.0002812

d_all_long <- pivot_longer(d_all, PIside_tray:PIside_floor)

ggplot(subset(d_all_long, condition=='pentadecane'), aes(x = name, y = value, fill = name)) +
  geom_boxplot(notch = T) +
  scale_fill_viridis(discrete = T, alpha = 0.6) +
  geom_jitter(color = 'black', width = 0.2, size = 0.4) +
  ylim(-1,1) +
  coord_flip() +
  theme_classic() +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey')
```

Stimulus side has a significant effect in the anova. The Kolmogorov Smirnov test
shows that the distributions are significantly different, and the Wilcoxon rank
sum test supports a difference in the medians, too.

BUT this only holds for larvae deposited on the floor, NOT those that were
deposited on the trays. Floor larvipositions seems to be governed by different
cues!

```{r}
Anova(aov(PItray ~ stim.right + dayday, data=subset(d_all, condition=='pentadecane'))) # very weak effect of day
ks.test(subset(d_all, condition=='pentadecane' & stim.right=='stim')$PItray,
        subset(d_all, condition=='pentadecane' & stim.right=='con')$PItray) # p = 0.09955
wilcox.test(subset(d_all, condition=='pentadecane' & stim.right=='stim')$PItray,
        subset(d_all, condition=='pentadecane' & stim.right=='con')$PItray) # p = 0.09955

ks.test(subset(d_all, condition=='pentadecane' & stim.right=='stim')$PIside_tray,
        subset(d_all, condition=='pentadecane' & stim.right=='con')$PIside_tray)

ks.test(subset(d_all, condition=='pentadecane' & stim.right=='stim')$PIside_floor,
        subset(d_all, condition=='pentadecane' & stim.right=='stim')$PIside_floor)

ggplot(subset(d_all_long, condition=='pentadecane'), aes(x = name, y = value, fill = stim.right)) +
  geom_boxplot(notch = T) +
  scale_fill_viridis(discrete = T, alpha = 0.6) +
  ylim(-1,1) +
  coord_flip() +
  theme_classic() +
  geom_point(position = position_jitterdodge(jitter.width = 0.15), size = 0.5) +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'grey')
```