---
title: "brms_analysis_summary"
output:
  pdf_document: default
date: "2022-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bayesian model for female larviposition choice analysis.

Import libraries:
```{r, echo = FALSE}
library('tidyr')
library('purrr')
library('brms')
library('dplyr')
```

## Load the data

```{r}
d_sub <- read.csv('/Users/addena/Documents/R/LSTM/Sand_vs_leaflitter_August.csv', 
                  sep=",") # substrate experiment
d_nat <- read.csv('/Users/addena/Documents/R/LSTM/naturalistic_exp.csv', 
                  sep=",") # naturalistic experiment
d_pent <- read.csv('/Users/addena/Documents/R/LSTM/C15_vs_PO_data.csv', 
                  sep=",") # pentadecane experiment
```

## Condense the data into one data frame

We remove cage 7 from the 'naturalistic' condition as it was terminated 
prematurely.

```{r}
d_sub <- aggregate(cbind(sand, leaf) ~ cage, data = d_sub, FUN = sum, na.rm = T)
d_sub$experiment <- c(rep('3_substrate', 10))
d_sub <- rename(d_sub, stim = leaf)
d_sub <- rename(d_sub, con = sand)

d_nat <- aggregate(cbind(stim, con) ~ cage, data = d_nat, FUN = sum, na.rm = T)
d_nat$experiment <- c(rep('1_naturalistic', 12))
d_nat <- d_nat[-7,] # remove cage 7 (terminated prematurely)

d_pent <- aggregate(cbind(stim, con) ~ cage, data = d_pent, FUN = sum, na.rm = T)
d_pent$experiment <- c(rep('2_pentadecane', 12))

data <- rbind(d_sub, d_nat, d_pent)
data
```
## Turn the data into binary format

Each pupa in the 'stim' [stimulus] category counts as a success ('1') and each 
pupa in the 'con' [control] category counts as a failure ('0'). Stimulus and 
control are defined as follows for the three experimental conditions:

Substrate: Stimulus = leaf litter, Control = sand; 
Pentadecane: Stimulus = pentadecane, Control = paraffin oil; 
Naturalistic: Stimulus = conditioned sand (20 pupae), 
              Control = unconditioned sand

```{r}
data |> mutate(binary = map2(stim, con, 
                                   ~ c(rep(1, .x), 
                                       rep(0, .y)))) -> data_binary
data_binary <- unnest(data=data_binary, cols=binary)
class(data_binary$cage) <- 'factorial'
class(data_binary$experiment) <- 'factorial'
data_binary
```
## Prepare brms model

with 'cage' as group-level effect and 'experiment' as 
population-level effect. Adjust beta priors from flat to normal.

```{r}
formula0 <- brms::brmsformula('binary ~ experiment + (1|cage)')
prior_b <- prior(normal(0,5), class = b)
formula0
prior_b
```

As the dependent variable is binary, samples will be drawn from a Bernoulli 
distribution with logit link.

```{r, echo=FALSE}
data_m0 <- brms::brm(data = data_binary, formula = formula0, 
                     family = bernoulli(link='logit'), prior = prior_b)
data_m0
```

## Evaluate model

Rhat and Bulk_ESS values suggest that the model converged well.

To confirm, we plot the model to visually inspect the chains.

```{r}
plot(data_m0)
```
Chains look good, the model has converged well. 

## Model results

We can now look at conditional effects predicted by the model.

```{r}
p <- conditional_effects(data_m0)
plot(p)
p[[1]]
```
The model supports a preference only in the 'substrate' condition, with an 
estimated preference of 0.9067 for '1', i.e. leaf litter. No preference is 
supported for the 'pentadecane' and 'naturalistic' experiments. Thus, the 
conditional effects support the results from behavioural experiments and mirror
the preference index analysis.

## Hypothesis tests

Just to make sure, we can check the outcomes for each individual population-
level effect:

```{r}
hyp1 <- hypothesis(data_m0, 'Intercept > 0')
plot(hyp1)
hyp1
```

Intercept is a dummy variable for the 'naturalistic' condition. The distribution
has an estimated median of 0.06 and does not support a preference for either 
side.

```{r}
hyp2 <- hypothesis(data_m0, 'experiment2_pentadecane > 0')
plot(hyp2)
hyp2
```

The distribution for the 'pentadecane' condition has an estimate of -0.06 and 
does not support a preference for either side.

```{r}
hyp3 <- hypothesis(data_m0, 'experiment3_substrate > 0')
plot(hyp3)
hyp3
```

The distribution for the 'substrate' condition clusters around an estimate of 
2.22 with an infinite evidence ratio and a posterior probability of 1, 
supporting a strong preference for choosing '1', i.e. leaf litter.